# Shiny Basics - code

## Useful links

* http://shiny.rstudio.com/gallery
* http://www.showmeshiny.com/
* http://shiny.rstudio.com/tutorial/
* http://shiny.rstudio.com/articles/
* http://shiny.rstudio.com/gallery/widget-gallery.html
* https://www.htmlwidgets.org/

## Course package data 
```{r message=FALSE}
library(shiny)
data(movies, package="jrYr")
```

## Input: control widgets
### Select box

```{r, eval = FALSE}
selectInput(inputId = "movie_type", # unique id 
            label = "Movie genre", # Text for web
            choices = c("Romance", "Action", "Animation"), 
            selected = "Action")
```

### Slider 

```{r, eval = FALSE}
sliderInput(inputId = "movie_rating", 
            label = "Movie rating", 
            min = 0, max = 10, value = 5)
```

### Numeric input

```{r, eval = FALSE}
numericInput(inputId = "movie_length", 
          label = "Movie length", 
          min = 1, max = 5220, value = 100)
```

## render functions
### Example: movies

```{r, eval = FALSE}
renderText(input$movie_type)
```

### Example: movies

```{r eval=FALSE}
renderText({
  type = movies[,input$movie_type] == 1
  nrow(movies[type,])
})
```
Or
```{r eval=FALSE}
renderPlot({
  type = movies[,input$movie_type] == 1
  hist(movies[type,]$length)
})
```

# Reactive programming
## Problem

This is inefficient

```{r eval=FALSE}
renderText({
  type = movies[,input$movie_type] == 1
  nrow(movies[type,])
})
renderPlot({
  type = movies[,input$movie_type] == 1
  hist(movies[type,]$length)
})
```

## Solution: `reactive()`

```{r eval=FALSE}
sub_movies = reactive({
  type = movies[,input$movie_type] == 1
  movies[type,]
})
```
Gives

```{r eval=FALSE}
renderText(nrow(sub_movies()))
renderPlot(hist(sub_movies()$length))
```

## `eventReactive()`
### Problem

We may not want the app to update every time we change an input

```{r eval=FALSE}
selectInput("movie_genre", label = "Movie genre", 
            c("Romance", "Action", "Animation"))

sub_movies = reactive({
  type = movies[,input$movie_type] == 1
  movies[type,]
})
#??
```

### Solution: `actionButton`, `actionLink` * `eventReactive()`

```{r, eval = FALSE}
selectInput("movie_genre", label = "Movie genre", 
            c("Romance", "Action", "Animation"))
actionButton("plot", "Plot it now!!")
```

```{r, eval = FALSE}
sub_movies1 = eventReactive(input$plot,{
  type = movies[,input$movie_genre] == 1
  movies[type,]
})
renderPlot(hist(sub_movies1()$length))
```

## Solution: `reactiveValues` and `observeEvent`

```{r eval=FALSE}
rvs = reactiveValues(data=movies)
```

```{r eval=FALSE}
observeEvent(input$romance, {
    type = movies[, "Romance"] == 1
    rvs$data = movies[type, ]
})
observeEvent(input$action, {
    type = movies[, "Action"] == 1
    rvs$data = movies[type, ]
})
renderPlot(hist(rvs$data[, "length"]))
```

## The `isolate` function
### Problem

```{r, eval = FALSE}
sliderInput("n", "Sample size", 10, 500, 100)
actionButton("romance", "Romance")
actionButton("action", "Action")
```


```{r eval=FALSE}
observeEvent(input$romance, {
    m =  movies[movies[, "Romance"] == 1, ]
    rows = sample(1:nrow(m), input$n)
    rvs$data = m[rows, ]
})
# Similar for Action
renderPlot(hist(rvs$data[, "length"], 
                main=paste("Sample size:", input$n)))
```

### Solution: `isolate`

```{r eval=FALSE}
renderPlot(hist(rvs$data[, "length"], 
                main=paste("Sample size:", isolate(input$n))))
```

## The `observe()` function
```{r, eval = FALSE, results="hide"}
data(movies, package = "jrShiny")
selectInput("movie_type", label = "Movie genre", 
            c("Romance", "Action", "Animation"))
rvs = reactiveValues(data = movies, sentence = NULL)
observe({
  rvs$data = movies[movies[, input$movie_type] == 1, ]
})

renderPlot(hist(rvs$data[, "length"]))
```
Doesn't really make sense to use an `observe()` here, just illustrating what it does.

# htmlwidgets
## The **plotly** package 

```{r, out.width="90%", message=FALSE, warning = FALSE}
library("ggplot2")
data(movies, package = "jrShiny")
g = ggplot(movies, aes(x = length, y = rating)) + 
  geom_point()
g
```

```{r, message=FALSE, warning = FALSE}
library("plotly")
ggplotly(g)
```

## Shiny and **plotly**

```{r, eval = FALSE, message=FALSE, warning = FALSE}
library("ggplot2")
library("plotly")
data(movies, package = "jrShiny")
selectInput("movie_type", label = "Movie genre",
            c("Romance", "Action", "Animation"))
rvs = reactiveValues(data = movies)

observeEvent(input$movie_type, {
  rvs$data = movies[movies[, input$movie_type] == 1, ]
})
renderPlotly({
  ggplotly(ggplot(rvs$data, aes(x = length)) +
    geom_histogram())
})
```

## Interactive tables with DT

```{r, message=FALSE, warning = FALSE}
library("DT")
datatable(movies)
```

## Shiny and **DT**

```{r, eval = FALSE, message=FALSE, warning = FALSE}
library("ggplot2")
library("plotly")
library("DT")
data(movies, package = "jrShiny")
selectInput("movie_type", label = "Movie genre",
            c("Romance", "Action", "Animation"))
rvs = reactiveValues(data = movies)
observeEvent(input$movie_type, {
  rvs$data = movies[movies[, input$movie_type] == 1, ]
})
renderPlotly({
  ggplotly(ggplot(rvs$data, aes(x = length)) +
    geom_histogram())
})
renderDataTable({
  datatable(rvs$data[,1:5],
            options = list(lengthMenu = c(5, 30, 50),
                           pageLength = 5))
})
```

# A shiny app
## The `ui.R` file

```{r eval=FALSE, results="hide", cache=FALSE, message=FALSE}
library("shiny")
fluidPage(
  titlePanel("Shiny happy people"), #title
  ## Sidebar with a slider input for no. of points
  sidebarLayout( 
    sidebarPanel(
      sliderInput("n", "Number of points:", min = 1, 
                                  max = 50, value = 30)
    ),
    ## Show a plot of the generated distribution
    mainPanel(plotOutput("scatter"))
  )
)
```  

## The `server.R` file

```{r eval = FALSE, results="hide", message=FALSE}
library("shiny")
# Function always has input & output
function(input, output) {
  # Expression that generates a plot. 
  # A call to renderPlot indicates that:
  #  1) It is "reactive" and therefore should 
  #   re-execute automatically when inputs change
  #  2) Its output type is a plot
  output$scatter = renderPlot({plot(rnorm(input$n))})
}
``` 

## Output objects

Output function | creates 
----------------|---------
`htmlOutput` | raw HTML
`imageOutput` | image
`plotOutput` | plot
`tableOutput` | table
`textOutput` | text
`uiOutput` | raw HTML
`verbatimTextOutput` | text

## The `fluidPage` function

```{r eval = FALSE}
fluidPage(
  titlePanel("Title panel"),# Title
  ## Sidebar style
  sidebarLayout(
    sidebarPanel("The sidebar"),
    mainPanel("Main panel")
  )
)
``` 

## `sidebarLayout`: swap sides

```{r eval = FALSE}
sidebarLayout(position="right",
  sidebarPanel("The sidebar"),
  mainPanel("Main panel")
)
```

## `sidebarLayout`

```{r eval = FALSE}
sidebarLayout(
  sidebarPanel("The sidebar", p("Choose an option")),
  mainPanel("Main panel")
)
```   

## Layout

```{r eval = FALSE}
ui = fluidPage(
  titlePanel("I love movies"), #Title
  fluidRow( # Define a row
    column(4, # Two columns: width 4 & 8
           wellPanel(
             selectInput("movie_type", label = "Movie genre", 
                         c("Romance", "Action", "Animation"))
           )       
    ),
    column(8, plotOutput("scatter"))
  )
)
```

## Tabset layouts

```{r, eval=FALSE}
mainPanel(
  tabsetPanel(type = "tabs", 
    tabPanel("Plot", plotOutput("plot")), 
    tabPanel("Summary", verbatimTextOutput("summary")), 
    tabPanel("Table", tableOutput("table"))
  )
)
``` 

# **shinydashboard**
## The **shinydashboard** package

```{r, eval = FALSE}
dashboardPage(
  dashboardHeader(title = "My first app"),
  dashboardSidebar(), 
  dashboardBody()
)
```

## The **shinydashboard** package

```{r, eval = FALSE}
dashboardPage(
  dashboardHeader(title = "My first app"),
  dashboardSidebar(
    sidebarMenu(
      menuItem("Page 1"), 
      menuItem("Page 2")
    )
  ), 
  dashboardBody()
)
```

## The **shinydashboard** package

```{r, eval = FALSE}
dashboardPage(
  dashboardHeader(title = "My first app"),
  dashboardSidebar(
    sidebarMenu(
      menuItem("Page 1", tabName = "p1"), 
      menuItem("Page 2", tabName = "p2")
    )
  ), 
  dashboardBody(
    tabItems(
      tabItem("p1", h2("This is page 1")), 
      tabItem("p2", h2("This is page 2"))
    )
  )
)
```

## The **shinydashboard** package

```{r, eval = FALSE}
dashboardPage(
  dashboardHeader(title = "My first app"),
  dashboardSidebar(
    sidebarMenu(
      menuItem("Page 1", tabName = "p1"), 
      menuItem("Page 2", 
               menuSubItem("Sub page 1", tabName = "sp1"), 
               menuSubItem("Sub page 2", tabName = "sp2"))
    )
  ), 
  dashboardBody(
    tabItems(
      tabItem("p1", h2("This is page 1")), 
      tabItem("sp1", h2("This is page 2, sub page 1")), 
      tabItem("sp2", h2("This is page 2, sub page 2"))
    )
  )
)
```

# Row layout
## Example Row layout

```{r, eval = FALSE}
dashboardBody(
  fluidRow(
    box(plotOutput("plot"), 
        title = "This is my plot", 
        width = 8), 
    box(sliderInput("n", "slider", 
                    min = 1, max = 100, value = 10), 
        title = "This is my slider", 
        width = 4)
  )
)
```

