```{r, message=FALSE}
library(shiny)
data(movies, package="jrYr")
```

# Introduction to Shiny
## Shiny

* https://goo.gl/tzZs3T
* http://shiny.rstudio.com/gallery
* http://www.showmeshiny.com/
* http://shiny.rstudio.com/tutorial/
* http://shiny.rstudio.com/articles/

## Task

Create a new R markdown document

`File` -> `New File` -> `R markdown`


## Code chunks

```` ```{r} ````     
`      # R code`     
```` ``` ````

* Everything between the back ticks is evaluated as R code

## Chunk options
## echo

* For hiding a code chunk 

```` ```{r, echo = FALSE} ````     
`      # R code`     
```` ``` ````

## eval

* Whether or not to actually run the code

```` ```{r, eval = FALSE} ````     
`      # R code`     
```` ``` ````

## Combining

* Of course you can combine as many options as you want
  
```` ```{r, echo = FALSE, eval = FALSE} ````     
`      # R code`     
```` ``` ````

## Shiny in R markdown

* We can use Shiny functionality within an Rmarkdown file

* Just create an R markdown file with the header

\noindent`---`    
`title: "First app"`    
`runtime: shiny`    
`---` 


# Input: control widgets
## Input: control widgets

## Select box

```{r, eval = FALSE, cache=FALSE}
selectInput(inputId = "movie_type", # unique id 
            label = "Movie genre", # Text for web
            choices = c("romance", "action", "thriller"), 
            selected = "action")
```

## Slider 

```{r, eval = FALSE, cache=FALSE}
sliderInput(inputId = "movie_rating", 
            label = "Movie rating", 
            min = 0, max = 10, value = 5)
```

## Numeric input

```{r, eval = FALSE, cache=FALSE}
numericInput(inputId = "movie_length", 
          label = "Movie length", 
          min = 1, max = 330, value = 100)
```

## Widgets

1)  http://shiny.rstudio.com/gallery/widget-gallery.html


# Rendered outputs
## Rendered output

Function | Output type
---------|------------
`renderPlot()` | R graphics output
`renderPrint()` | printed output
`renderTable()`	| Data frame, matrix
`renderText()` | Character vectors

## Example: movies

```{r eval=FALSE, cache=FALSE}
renderText(input$movie_type)
```

## Example: movies

* If we want more than one line of code in the render function, we need `{}`

```{r eval=FALSE, cache=FALSE}
renderText({
  type = movies[,input$movie_type] == 1
  nrow(movies[type,])
})
```
Or
```{r eval=FALSE, cache=FALSE}
renderPlot({
  type = movies[,input$movie_type] == 1
  hist(movies[type,]$length)
})
```

## Task

In the previous task we created a `selectInput()` to let the user choose a movie classification
```{r, eval = FALSE, cache=FALSE}
selectInput(inputId = "movie_type",
            label = "Movie classification",
            choices = c("U", "PG", "12A", "15", "18"), 
            selected = "12A")
```

- Using `renderText()`, display the number of films within each `classification`. For example, the number of "PG" films 
```{r,eval = FALSE, cache=FALSE}
data(movies, package = "jrYr")
pg_movies = movies[movies$classification == "PG",]
nrow(pg_movies)
```

# Reactive programming
## Problem

This is inefficient

```{r eval=FALSE, cache=FALSE}
renderText({
  type = movies[,input$movie_type] == 1
  nrow(movies[type,])
})

renderPlot({
  type = movies[,input$movie_type] == 1
  hist(movies[type,]$length)
})
```

## Solution: `reactiveValues()` 

```{r eval=FALSE, cache=FALSE}
rvs = reactiveValues(data = movies)
```

## Solution: `observe()`

```{r, eval = FALSE, cache=FALSE}
observe({
    type = movies[,input$movie_type] == 1
    rvs$data = movies[type,]
})
```

## Solution

```{r, eval = FALSE, cache=FALSE}
rvs = reactiveValues(data = movies)

observe({
  type = movies[, input$movie_type] == 1
  rvs$data = movies[type, ]
})

renderText({
  nrow(rvs$data)
})

renderPlot({
  hist(rvs$data$length)
})
```

# Problem
## Problem

We may not want the app to update every time we change an input

## Solution: `observeEvent()`, `actionButton()`

```{r, eval = FALSE, cache=FALSE}
rvs = reactiveValues(data = movies)

actionButton("plot_button", "Plot it now!!!")

observeEvent(input$plot, {
  type = movies[, input$movie_type] == 1
  rvs$data = movies[type, ]
})

renderPlot({
  hist(rvs$data$length)
})
```

# htmlwidgets
## The **plotly** package 

```{r, eval = FALSE, cache=FALSE}
library("ggplot2")
data(movies, package = "jrYr")
g = ggplot(movies, aes(x = length, y = rating)) + 
  geom_point()
g
```

## The **plotly** package 

```{r, eval = FALSE, cache=FALSE}
library("plotly")
ggplotly(g)
```

## Shiny and **plotly**

```{r, eval = FALSE, cache=FALSE}
library("ggplot2")
library("plotly")
data(movies, package = "jrYr")
selectInput(inputId = "movie_type", 
            label = "Movie genre", 
            choices = c("romance", "action", "thriller"))
rvs = reactiveValues(data = movies)

observe({
  type = movies[, input$movie_type] == 1
  rvs$data = movies[type, ]
})

renderPlotly({
  ggplotly(ggplot(rvs$data, aes(x = length)) +
    geom_histogram())
})
```

## Interactive tables with DT

```{r, eval = FALSE, message=FALSE, warning = FALSE, cache=FALSE}
library("DT")
datatable(movies)
```

## Interactive tables with DT

```{r, eval = FALSE, results = "hide", message=FALSE, warning = FALSE, cache=FALSE}
library("ggplot2")
library("plotly")
library("DT")
data(movies, package = "jrYr")
selectInput(inputId = "movie_type",
            label = "Movie genre",
            choices = c("romance", "action", "thriller"))
rvs = reactiveValues(data = movies)
observeEvent(input$movie_type, {
  rvs$data = movies[movies[, input$movie_type] == 1, ]
})
renderDataTable({
  datatable(rvs$data[,1:5])
})
```

# A shiny app
## The `ui.R` file

```{r,eval = FALSE, results="hide", cache=FALSE, message=FALSE}
library("shiny")
fluidPage(
  titlePanel("Shiny happy people"), #title
  ## Sidebar with a slider input for no. of points
  sidebarLayout( 
    sidebarPanel(
      selectInput(inputId = "movie_type", # unique id 
            label = "Movie genre", # Text for web
            choices = c("romance", "action", "thriller"), 
            selected = "action")
    ),
    ## Show a plot of the generated distribution
    mainPanel(plotOutput("scatter"))
  )
)
```  

## The `server.R` file

```{r results="hide", message=FALSE,eval = FALSE, cache=FALSE}
library("shiny")
data(movies, package = "jrYr")

# Function always has input & output
function(input, output) {
  rvs = reactiveValues(data = movies)
  
  observeEvent(input$movie_type, {
    rvs$data = movies[movies[, input$movie_type] == 1,]
  })
  
  output$scatter = renderPlot({
    plot(x = rvs$data$duration, y = rvs$data$rating)
  })
  
}
```

## Output objects

Output function | creates 
----------------|---------
`htmlOutput` | raw HTML
`imageOutput` | image
`plotOutput` | plot
`tableOutput` | table
`textOutput` | text
`uiOutput` | raw HTML
`verbatimTextOutput` | text

## The `fluidPage` function

```{r results="hide", cache=FALSE,eval = FALSE}
fluidPage(
  titlePanel("Title panel"),# Title
  ## Sidebar style
  sidebarLayout(
    sidebarPanel("The sidebar"),
    mainPanel("Main panel")
  )
)
``` 

## `sidebarLayout`: swap sides

```{r results="hide", cache=FALSE,eval = FALSE}
sidebarLayout(position="right",
  sidebarPanel("The sidebar"),
  mainPanel("Main panel")
)
```

## `sidebarLayout`

```{r results="hide",eval = FALSE, cache=FALSE}
sidebarLayout(
  sidebarPanel("The sidebar", p("Choose an option")),
  mainPanel("Main panel")
)
```   

## Layout

```{r cache=FALSE,eval = FALSE}
ui = fluidPage(
  titlePanel("I love movies"), #Title
  fluidRow( # Define a row
    column(4, # Two columns: width 4 & 8
           wellPanel(
             selectInput("movie_type", label = "Movie genre", 
                         c("Romance", "Action", "Animation"))
           )       
    ),
    column(8, plotOutput("scatter"))
  )
)
```
